---
title: Client Side Calculations Based on Graphics Interaction
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo    = TRUE,
	message = FALSE,
	warning = FALSE
)
library(quantmod)
library(tidyverse)
library(plotly)
library(reactable)
library(sparkline)
```

## Interactive Stock Price Plot

This graph displays the development of some adjusted stock prices. 
Note that you can toggle the visibility of the ``code chunks`` that generate the data and visualization.

```{r stockPrices}
# Get Adjusted Stock Prices and
symbols <- c("AAPL", "MSFT", "SPY")
invisible(quantmod::getSymbols(symbols, 
                     from = "2021-11-30", to = "2021-12-31",
                     auto.assign = TRUE, warnings = FALSE))

prices_xts <- 
  xts::merge.xts(AAPL$AAPL.Adjusted, MSFT$MSFT.Adjusted, SPY$SPY.Adjusted)

## export csv for separate upload
# zoo::write.zoo(prices_xts, 
#           file = file.path("H:", "Temp", "prices.csv"),
#           quote = FALSE, sep = ",")

prices_long <- 
  prices_xts %>% 
  tibble::as_tibble(rownames = "date") %>% 
  dplyr::rename(Apple = AAPL.Adjusted, Microsoft = MSFT.Adjusted, SandP500 = SPY.Adjusted) %>% 
  dplyr::mutate(date = as.Date(date)) %>% 
  tidyr::gather(key = "asset", value = "idx", -date) %>% 
  dplyr::group_by(asset) %>% 
  dplyr::arrange(asset, date) 

plt_prices <- 
  plot_ly(prices_long, x = ~date, y = ~idx, color = ~asset,
          type = "scatter", mode = "lines") 

plt_prices
```

Note the interactivity of the plot. 
In particular the time series to be displayed can be selected by clicking on the corresponding entry in the legend. 
One can also zoom in and out via the ``navigation bar`` or by clicking and dragging horizontally.
The full period is restored by double-clicking into the plot or pressing ``auto-scale`` in the navigation bar.

## Corresponding Stock Return Summary Table

The following table displays some summary statistics and visuatization of the distribution for all stocks returns over the full time period. 

```{r summaryTable}
# Return info in long table
return_long <-
  prices_long %>% 
  dplyr::mutate(., sRet = idx / dplyr::lag(idx) - 1)

data <- 
  return_long %>% 
  dplyr::select(asset, sRet) %>% 
  dplyr::group_by(asset) %>% 
  dplyr::summarize(mean = mean(sRet, na.rm = TRUE), 
                   sd   = sd(sRet, na.rm = TRUE),
                   sRet = list(sRet))  %>% 
  mutate(box = NA, sparkline = NA)
  
rtbl_returns <- 
  reactable(data = dplyr::select(data, 
                                 asset, mean, sd, box, sparkline), 
          columns = list(
            mean = colDef(format = colFormat(percent = TRUE, digits = 3)),
            sd   = colDef(format = colFormat(percent = TRUE, digits = 3)),
            box       = colDef(cell = function(value, index) {
              sparkline(data$sRet[[index]], type = "box")
            }),
            sparkline = colDef(cell = function(value, index) {
              sparkline(data$sRet[[index]])
            })
))

rtbl_returns
```

## Recalculate Summary Table Based on Graphics Interaction <!-- {.tabset} -->

The challange is to recalculate the table such that it 

- only displays the assets selected in the plot,
- calculates the statistics only for the visible time period,
- all calculations take place in the browser, no Shiny!

The following plot is without the Microsoft stock and only for the sub period from December 10, 2021 to December 24.

### Filterd Plot

```{r combinedStatic}
prices_long_filtered <- 
  prices_long %>% 
  dplyr::filter(date >= as.Date("2021-12-10") & date <= as.Date("2021-12-24")) %>% 
  dplyr::filter(asset != "Microsoft")

plt_prices_filtered <- 
  plot_ly(prices_long_filtered, x = ~date, y = ~idx, color = ~asset,
          type = "scatter", mode = "lines") 
plt_prices_filtered
```

### Desired Table Based on Filter

The desired outcome is a revised table reflecting the user interaction in the plot.

```{r}
return_long_filtered <-
  prices_long %>% 
  dplyr::mutate(., sRet = idx / dplyr::lag(idx) - 1) %>% 
  dplyr::filter(date >= as.Date("2021-12-10") & date <= as.Date("2021-12-24")) %>% 
  dplyr::filter(asset != "Microsoft")

data_filtered <- 
  return_long_filtered %>% 
  dplyr::select(asset, sRet) %>% 
  dplyr::group_by(asset) %>% 
  dplyr::summarize(mean = mean(sRet, na.rm = TRUE), 
                   sd   = sd(sRet, na.rm = TRUE),
                   sRet = list(sRet))  %>% 
  mutate(box = NA, sparkline = NA)
  
rtbl_returns_filtered <- 
  reactable(data = dplyr::select(data_filtered, 
                                          asset, mean, sd, box, sparkline), 
          columns = list(
            mean = colDef(format = colFormat(percent = TRUE, digits = 3)),
            sd   = colDef(format = colFormat(percent = TRUE, digits = 3)),
            box       = colDef(cell = function(value, index) {
              sparkline(data_filtered$sRet[[index]], type = "box")
            }),
            sparkline = colDef(cell = function(value, index) {
              sparkline(data_filtered$sRet[[index]])
            })
))
rtbl_returns_filtered
```

## Implementation Ideas

The initial focus of this project is to effectively organise the data and do the calculationsin JavaScript

- use some specific data structure and calculation routines from [tidy.js](https://pbeshai.github.io/tidy/), [arquero.js](https://uwdata.github.io/arquero/), [danfo.js](https://danfo.jsdata.org/), or other?
- Can [Hal9](https://hal9.com/) be an integrating framework?
- Ideally, only the stock prices data is initially provided - everything else, including returns should be done in the browser.
- The status of the plotly object can be retrieved with some attributes. The following seem particularly relevant:
  - The selected starting point of the x-axis in plotly is stored in ``el.layout.xaxis.range[0]``. 
  - The visibility of a trace can be checked with the attribute [visible](https://plotly.com/javascript/reference/scatter/#scatter-visible).
- There will be more complex transformations and calculations (regressions, annualalisations).

Eventually, the framework should be an [htmlwidget](http://www.htmlwidgets.org/) R-Package.

